### 2012-10-11 1.tex
Completion of draft of chapter 3

\section{\emph{Táxis} and generativity}
%\emph{τάξις}

(\emph{táxis} `order, arrangement'; the prefix \emph{syn-} means `together')

This etymology is almost completley uninformative about the modern study of syntax. 
The notion of derivation, most associated with the transformational grammars of Zellig Harris and Noam Chomsky

This etymology gives little clue to just what syntacticians do, however. This is not merely a consequence of the ``generative revolution'', as the predecessors of transformational grammars associated with Zellig Harris and Noam Chomsky were also more than theories of arrangement, but also instantiate the notion of derivation.

@@

@@

There are very few investigating infants' knowledge of phonological alternations. \citet{White2008} expose 8.5-month-old and 10-month-old infants to an artificial language in which the voicing of fricatives is contrastive, but voiced and voiceless variants of stops are in a complementary distribution, appearing only after vowels (\emph{na-bevi}) and after voiceless consonants (\emph{rot-pevi}), respectively. After familiarization, infants at both ages prefer to listen to stimuli which preserve the complementary distribution over those which disrupt it (e.g., \emph{na poli}, \emph{rot boli}). As \citeauthor{White2008} note, this suggests that infants have grouped \emph{pevi} and \emph{bevi} together and extracted the stop voicing alternation from the artifical language.

@@

As first observed by \citet{Darwin1877}, the first words infants recognize are names, their own and those of caretakers. Recent studies suggest that infants learn these names as early as 4 months of age \citep{Bortfeld2005,Mandel1995,Tincoff1999}. Infants as young as 6 months of age have learned the meaning of familiar words \citep{Bergelson2012}. By 8 months of age, infants are able to locate familiar and novel words in utterances \citep{Jusczyk1997,Seidl2006}.
\section{Outline of the dissertation}

The remainder of the dissertation consists of three case studies which provide support for the novel and contentious predictions of the minimal phonotactic model sketched above

Chapter \ref{gradience} considers evidence from wordlikeness rating tasks. It is argued that intermediate well-formedness ratings are obtained whether or not the categories in question are graded. A primitive categorical model of wordlikeness using prosodic representations is outlined, and shown to predict English speakers' gradient wordlikeness judgements as well as state-of-the-art gradient wellformedness models. The gradient models do not contribute any additional coverage once categorical wellformedness is accounted for.

Chapter \ref{turkish} considers the relationship between lexical generalizations, phonological alternations, and speakers' nonce word judgements with a focus on Turkish vowel patterns. It is shown that even exception-filled phonological generalizations have a robust effect on wellformedness judgements, but that statistically reliable phonotactic generalizations go unlearned when they are not derived from phonological alternations.
%Some arguments against Becker et al.

Chapter \ref{clusters} investigates the role of phonological alternations in constraining lexical entries, focusing specifically on medial consonant clusters in English. Static phonotactic constraints previously proposed to describe gaps in the inventory of medial clusters are shown to be statistically unsound. On the other hand, phonological alternations impose robust restrictions on the cluster inventory. The remaining gaps in the cluster inventory are due to the sparse nature of the lexicon, not static phonotactic restrictions.

%% NOTES HERE
%a. & \emph{B}[uːx]          & \emph{B}[yːç]\emph{er} & `book(s)' \\
%   & \emph{L}[ɔx]           & \emph{L}[øç]\emph{er}  & `hole(s)' \\
%   &\emph{T}[ɔx]\emph{ter} & \emph{T}[øç]\emph{ter}
%   & \emph{N}[ax]\emph{t}   & \emph{N}[ɛç]\emph{te} & `night' \\
%
% While the bulk of evidence for detailed phonological entries is found in infants over a year of age \citep[e.g.,]{Werker2002,Fennell2003,Fennell2006,Stager1997,Swingley2000,Zamuner2006},
%9-month-old infants already have learned the predominant stress pattern of their language \citep{Jusczyk1993a}.

>>>>

One question is: why bother?

It is certainly possible to imagine otherwise. \citet[][320]{Zimmer1969} writes that phonotactic generalizations ``have no observable consequences in the course of the normal use of the language'', and \citeauthor{PE} echo this sentiment more recently:

\begin{quote}
Even if we, as linguists, find some generalizations in our description of the lexicon, there is no reason to posit these generalizations as part of the speaker's knowledge of their language, since they are computationally inert and thus irrelevant to the input-output mapping that the grammar is responsible for. \citep[][18]{PE}
\end{quote}

This represents a principled null hypothesis, but can be quickly dismissed in light of speaker's ready judgements of possible and impossible words observed by \citeauthor{Chomsky1965}. Metalinguistic judgements of these sorts, reviewed in Chapter \ref{wordlikeness}, are not the only task which exemplify phonotactic knowledge; infants and adults are also thought to make use \emph{possible word constraint} in more quotidien tasks like recognizing words in running speech \citep[e.g.,][]{Brown1956,Mattys1999,McQueen1998b,Norris1997}.

@@

This is precisely the case for

@@

Very few studies have investigated young infants' knowledge of phonological alternations. One exception is a fascinating study by \citet{White2008}. Simplifying somewhat, the experimenters expose 8.5-month-old and 10-month-old infants to an artificial language in which the voicing of fricatives is contrastive, but voiced and voiceless variants of plosives are in a complementary distribution, appearing only after vowels (\emph{na-bevi}) and after voiceless consonants (\emph{rot-pevi}), respectively. After familiarization, infants at both ages prefer to listen to stimuli which preserve the complementary distribution over those which disrupt it (e.g., \emph{na-poli}, \emph{rot-boli}). As \citeauthor{White2008} note, this suggests that infants have both extracted the plosive voicing alternation and grouped \emph{pevi} and \emph{bevi} together.

@@

As first observed by \citet{Darwin1877}, the first words infants recognize are names, their own and those of caretakers. Recent studies suggest that infants learn these names as early as 4 months of age \citep{Bortfeld2005,Mandel1995,Tincoff1999}. Infants as young as 6 months of age have learned the meaning of familiar words \citep{Bergelson2012}. By 8 months of age, infants are able to locate familiar and novel words in utterances \citep{Jusczyk1997,Seidl2006}.

\section{Outline of the dissertation}

The remainder of the dissertation consists of three case studies which provide support for the novel and contentious predictions of the minimal phonotactic model sketched above.

Chapter \ref{gradience} considers evidence from wordlikeness rating tasks. It is argued that intermediate well-formedness ratings are obtained whether or not the categories in question are graded. A primitive categorical model of wordlikeness using prosodic representations is outlined, and shown to predict English speakers' wordlikeness judgements at least as acccurately as state-of-the-art gradient wellformedness models. Once categorical effects are controlled for, gradient models are uncorrelated with well-formedness ratings.

Chapter \ref{turkish} considers the relationship between lexical generalizations, phonological alternations, and speakers' nonce word judgements with a focus on Turkish vowel patterns. It is shown that even exception-filled phonological generalizations have a robust effect on wellformedness judgements, but that statistically reliable phonotactic generalizations go unlearned when they are not derived from phonological alternations.

Chapter \ref{clusters} investigates the role of phonological alternations in constraining lexical entries, focusing specifically on medial consonant clusters in English. Static phonotactic constraints previously proposed to describe gaps in the inventory of medial clusters are shown to be statistically unsound, whereas phonological alternations impose robust restrictions on the cluster inventory. The remaining gaps in the cluster inventory are attributed to the sparse nature of the lexicon, not static phonotactic restrictions.

%% NOTES HERE
%
%a. & \emph{B}[uːx]          & \emph{B}[yːç]\emph{er} & `book(s)' \\
%   & \emph{L}[ɔx]           & \emph{L}[øç]\emph{er}  & `hole(s)' \\
%   &\emph{T}[ɔx]\emph{ter} & \emph{T}[øç]\emph{ter}
%   & \emph{N}[ax]\emph{t}   & \emph{N}[ɛç]\emph{te} & `night' \\
%
% While the bulk of evidence for detailed phonological entries is found in infants over a year of age \citep[e.g.,]{Werker2002,Fennell2003,Fennell2006,Stager1997,Swingley2000,Zamuner2006},
%9-month-old infants already have learned the predominant stress pattern of their language \citep{Jusczyk1993a}.
%\emph{τάξις}

### 2012-10-11 2.tex
Completion of draft of chapter 3

%\citet{Saporta1958}
%\citet{Saporta1957}

@@

Despite the considerable attention given to the proposals of \emph{SPE} in the wake of that book's publication in 1968, the \emph{SPE} wordlikeness model has received almost no attention in the literature. At the risk of explaining what might be no more than an an accidental gap in the historical record, the novel aspects of the \emph{SPE} model---gradience derived from similarity to existing lexical entries---seem to have been overshadowed by compelling arguments against \citeauthor{SPE}'s assumption \citep[see also][]{Halle1962,Stanley1967} that wordlikeness contrasts derive solely from properties of underlying forms. \citet{Shibatani1973} observes that there are some generalizations about surface forms which give rise to wordlikeness contrasts, but cannot be stated as constraints on underlying forms. An example from German is shown in (\ref{fd}) below.

@@

\begin{quote}
Are there definitional concepts? Of course. For example, consider the concept \emph{odd number}. This seems to have a clear definition, a precise description\ldots{}No integer seems to sit on the fence, undecided as to whether it is quite even, or perhaps a bit odd. No odd number seems odder than any other odd number. \citep[274]{Armstrong1983}
\end{quote}

>>>>

%\citet{Fischer-Jorgensen1952}
%\citet{Saporta1958}
%\citet{Saporta1957}

@@

Despite the considerable attention given to the proposals of \emph{SPE} in the wake of that book's publication in 1968, the \emph{SPE} wordlikeness model has received almost no further attention in the literature. At the risk of explaining what might be no more than an an accidental gap in the literature, the novel aspects of \emph{SPE} model---gradience derived from similarity to existing lexical entries---may have been overshadowed by the many other contentious proposals in \emph{SPE}, and particularly by compelling arguments against the assumption that wordlikeness contrasts derive solely from properties of underlying forms. \citet{Shibatani1973} observes that there are some generalizations about surface forms which give rise to wordlikeness contrasts, but cannot be stated as constraints on underlying forms. An example from German is shown in (\ref{fd}) below.

@@

\begin{quote}
Are there definitional concepts? Of course. For example, consider the concept \emph{odd number}. This seems to have a clear definition, a precise description. [\ldots{}] No integer seems to sit on the fence, undecided as to whether it is quite even, or perhaps a bit odd. No odd number seems odder than any other odd number. \citep[274]{Armstrong1983}
\end{quote}
